---
title: "Developing the Deduplicating Process"
author: "Anna Ringwood"
date: "5/27/2021 (Last Updated: 5/28/2021)"
output: html_document
---

# **Cleaning and Deduplicating Bloomerang Constituent Data**

**Prior to beginning in R, complete the following steps:**

1. Download Bloomerang Constituents data (which will be an Excel Workbook by default)  
2. Open Bloomerang Constituents data in Excel  
3. Save Bloomerang Constituents data as a .csv file

* Call libraries `tidyverse`, `janitor`, and `lubridate` in R:

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

* Read Bloomerang Constituents CSV file into R as `bloomConstsRaw`, specifying `header = TRUE` and `na.strings = ""`:

```{r}
bloomConstsRaw <- read.csv("Z:/Emory QTM - Constituents - All Fields CSV.csv", header = TRUE, na.strings = "")
```

The resulting object is a data frame with 3,771 rows (records) and 78 columns (variables).

* Use `janitor`'s `clean_names()` to convert the variable names in `bloomConstsRaw` to `"lower_camel"`  
* Rename `todaySDate` to `todaysDate` and `viPsAndInfluencers` to `vipsAndInfluencers`  
* Coerce `createdDate` and `lastModifiedDate` to "POSIXct" class using `lubridate`'s `mdy_hm()`  
* Convert remaining date variables to "Date" class using `lubridate`'s `mdy()`:  
  * `birthdate`  
  * `firstTransactionDate`  
  * `largestTransactionDate`  
  * `latestTransactionDate`  
  * `secondTransactionDate`  
  * `todaysDate`
* Use `remove_str_make_num()` to remove "$" from and coerce the following variables to numeric type:   
  * `firstTransactionAmount`  
  * `largestTransactionAmount`  
  * `lastYearRaised`  
  * `lastYearRevenue`  
  * `latestTransactionAmount`  
  * `lifetimeRaised`  
  * `lifetimeRevenue`  
  * `secondTransactionAmount`  
  * `yearToDateRaised`  
  * `yearToDateRevenue`  
* Use `remove_str_make_num()` to remove "," from `numberOfTransactions` and coerce to numeric class  
* Convert all character variables' values to be completely lowercase using `stringr`'s `str_to_lower()`  
* Remove duplicate name variable `name1`

```{r}
### Define `remove_str_make_num()`:
remove_str_make_num <- function(df_column, char_to_remove){
  df_column <- as.numeric(str_remove(df_column, char_to_remove))
  return(df_column)
}
```

```{r warning = F}
bloomConstsClean <- bloomConstsRaw %>%
  clean_names("lower_camel") %>%
  rename(todaysDate = todaySDate, vipsAndInfluencers = viPsAndInfluencers) %>%
  mutate(createdDate = mdy_hm(createdDate), lastModifiedDate = mdy_hm(lastModifiedDate),
         across(c(contains("TransactionDate"), "birthdate", "todaysDate"), ~ mdy(.x)),
         across(contains(c("Amount", "Revenue", "Raised")), ~ remove_str_make_num(.x, "[$]")),
         numberOfTransactions = remove_str_make_num(numberOfTransactions, ","),
         across(where(is.character), str_to_lower)) %>%
  select(-name1)
```

*The resulting object is a data frame with 3,771 records and 77 variables.*

* Separate first row of aggregate totals from the individual records in the data frame:

```{r}
totalsBloomConsts <- bloomConstsClean[1,]
bloomConstsClean <- bloomConstsClean[-1,]
```

*Two objects result:*

*1. `totalsBloomConsts`: 1 record, 77 variables*  
*2. `bloomConstsClean`: 3770 records, 77 variables*

*The first object, `totalsBloomConsts`, is set aside and the remaining analysis concerns the 3770 individual records in the `bloomConstsClean` data frame.*

* Use `stringr`'s `str_replace_na()` to replace `NA` values with empty strings in name variables which will be used to create full name variables:  
  * `firstName`  
  * `middleName`  
  * `lastName`  
  * `nameTitle`  
  * `nameSuffix`
* Use `stringr`'s `str_replace_all()` to remove "." from name variables:  
  * `firstName`  
  * `middleName`  
  * `lastName`
* Use `stringr`'s `str_replace_all()` to replace "-" with " " in name variables:  
  * `firstName`  
  * `middleName`  
  * `lastName`
* Use `stringr`'s `str_squish()` to create a `fullName` variable from `firstName`, `middleName`, `lastName`, and `nameSuffix`
* Use `stringr`'s `str_squish()` to create a `longFullName` variable from `nameTitle`, `firstName`, `middleName`, `lastName`, and `nameSuffix`
* Revert the strings previously made empty back to `NA` values in name variables:  
  * `firstName`  
  * `middleName`  
  * `lastName`  
  * `nameTitle`  
  * `nameSuffix`

```{r}
bloomConstsClean <- bloomConstsClean %>%
  mutate(across(all_of(c("firstName", "middleName", "lastName", "nameSuffix", "nameTitle")),
                ~ str_replace_na(.x, " ")),
         across(all_of(c("firstName", "middleName", "lastName")), ~ str_replace_all(.x, "\\.", "")),
         across(all_of(c("firstName", "middleName", "lastName")), ~ str_replace_all(.x, "-", " ")),
         fullName = str_squish(str_c(firstName, middleName, lastName, nameSuffix, sep = " ")),
         longFullName = str_squish(str_c(nameTitle, firstName, middleName, lastName, nameSuffix, sep = " ")),
         across(all_of(c("firstName", "middleName", "lastName", "nameSuffix", "nameTitle")),
                ~ ifelse(.==" ", NA, as.character(.))))
```

*The fully cleaned data set, `bloomConstsClean`, contains 3770 records and 79 variables.*

# Segment Data

* Identify duplicate records based on `fullName`  
* Extract all records corresponding to the duplicate full names from `bloomConstsClean`  
* Identify the number of Individual and Organization account types associated with each duplicate name

```{r}
duplicateFullNamesSums <- bloomConstsClean %>%
  group_by(fullName) %>%
  summarize(nNames = n()) %>%
  filter(nNames != 1)

dupFullNameRecords <- bloomConstsClean %>%
  filter(fullName %in% pull(duplicateFullNamesSums, fullName))

numOrgIndAccts <- dupFullNameRecords %>%
  group_by(fullName) %>%
  summarize(nOrgs = sum(type == "organization"),
            nInds = sum(type == "individual"))
```

* Segment duplicate records based on the number of records categorized as "Individual" and "Organization"

```{r}
indOnlyRecords <- bloomConstsClean %>%
  filter(fullName %in% pull(filter(numOrgIndAccts, nOrgs == 0 & nInds != 0), fullName))
orgOnlyRecords <- bloomConstsClean %>%
  filter(fullName %in% pull(filter(numOrgIndAccts, nOrgs != 0 & nInds == 0), fullName))
bothTypesRecords <- bloomConstsClean %>%
  filter(fullName %in% pull(filter(numOrgIndAccts, nOrgs != 0 & nInds != 0), fullName))
```

Some scenarios to consider when collapsing duplicate records:

* Across all records for a name, the values for a variable are the same  
  * In this case, the value is copied over to the new record  
  
```{r}
identicalRecordVals <- function(df_column){
  identicalVals <- vector()
  for(i in 2:length(df_column)){
    identicalVals <- append(identicalVals, identical(df_column[i], df_column[i-1]))
  }
  if(sum(identicalVals == FALSE) == 0){
    verdict <- TRUE
  } else {
    verdict <- FALSE
  }
  return(verdict)
}
```

* Across all records for a name, there is at least one `NA` value in a variable but the remaining value(s) are the same  
  * In this case, the unique, non-NA value is copied over to the new record  
  
```{r}
oneUniqueField <- function(df_column){
  colNoNA <- na.omit(df_column)
  if(length(unique(colNoNA)) == 1){
    verdict <- TRUE
  } else {
    verdict <- FALSE
  }
  return(verdict)
}
```

* The values for a variable are additive and should be aggregated  
  * In this case, all values across the records are pasted together  
  
```{r}
aggregatableVars <- c("primaryEmailAddress", "accountNumber", "emailInterestType", "eventsAttended", "jobTitle", "relationshipToGeears", "vipsAndInfluencers")

aggregateVals <- function(df_column){
  aggregated <- ""
  for(row in 1:length(df_column)){
    if(row == length(df_column)){
      aggregated <- paste0(aggregated, df_column[row])
    }else{
      aggregated <- paste0(aggregated, df_column[row], "|")
    }
  }
  return(aggregated)
}
```

* Across all records for a name, there are at least two non-NA values for a variable  
  * In this case, a closer examination of the fields and values is required  
    * `engagementLevel`/`generosity`: highest value is kept  
    * `lastModifiedDate`: most recent date is kept  
    * `lastModifiedBy`: Whatever is attached to the kept `lastModifiedDate`
  
* Convert `generosity` and `engagementLevel` to integers so that the higher value can be kept

```{r}
indOnlyRecords$generosity <- as.numeric(as.character(factor(indOnlyRecords$generosity, levels = c("not scanned", "cold", "cool", "warm", "hot", "on fire!"), labels = c("0", "1", "2", "3", "4", "5"), ordered = TRUE)))
indOnlyRecords$engagementLevel <- as.numeric(as.character(factor(indOnlyRecords$engagementLevel, levels = c("cold", "cool", "warm", "hot", "on fire!"), labels = c("1", "2", "3", "4", "5"), ordered = TRUE)))

keepHighestVars <- c("engagementLevel", "generosity", "lastModifiedDate")

bloomConstsClean %>%
  select(contains("name")) %>%
  filter(firstName != informalName)

bloomConstsClean %>%
  group_by(lastModifiedBy, lastModifiedDate) %>%
  summarize(n())
```

* Combine the personal information across records

```{r}
collapseRecords <- function(dataset, full_name){
  
  # split transaction info and personal info
  transDF <- dataset %>%
  select(accountNumber, fullName, contains("transaction") | contains("revenue") | contains("raised"))
  personalInfoDF <- dataset %>%
  select(!contains(c("transaction", "revenue", "raised")),
         -c("envelopeName", "formalName", "recognitionName", "longFullName"))

  # create newRecord to hold deduplicated information
  newRecord <- bloomConstsClean[1,]
  newRecord[1,] <- NA
  
  # obtain all records corresponding to given name
  compareDF <- personalInfoDF %>%
    filter(fullName == full_name)
  
  # run identicalRecordVals() on all columns
  identicalVerdicts <- compareDF %>%
    summarize(across(everything(), ~ identicalRecordVals(.x)))
  # split variables based on True/False
  identicalTrue <- compareDF[,identicalVerdicts == TRUE, drop = FALSE]
  identicalTrueNames <- names(identicalTrue)
  identicalFalse <- compareDF[,identicalVerdicts == FALSE, drop = FALSE]
  identicalFalseNames <- names(identicalFalse)
  # add True results to newRecord
  newRecord[1, identicalTrueNames] <- compareDF[1, identicalTrueNames]

  # run oneUniqueField() on all remaining columns
  oneUniqueVerdicts <- identicalFalse %>%
    summarize(across(everything(), ~ oneUniqueField(.x)))
  # split based on True/False
  oneUniqueTrue <- identicalFalse[,oneUniqueVerdicts == TRUE, drop = FALSE]
  oneUniqueTrueNames <- names(oneUniqueTrue)
  oneUniqueFalse <- identicalFalse[,oneUniqueVerdicts == FALSE, drop = FALSE]
  oneUniqueFalseNames <- names(oneUniqueFalse)
  # add True results to newRecord
  uniqueValsInCol <- oneUniqueTrue %>%
    summarize(across(everything(), unique)) %>%
    na.omit()
  for(col in oneUniqueTrueNames){
    uniqueVal <- uniqueValsInCol[1, col]
    newRecord[1, col] <- uniqueVal
  }
  
  # run aggregateVals() on specific remaining columns
  aggregateResults <- oneUniqueFalse %>%
    summarize(across(any_of(aggregatableVars), ~ aggregateVals(.x)))
  # split based on what was/was not squished
  aggregatedTrue <- oneUniqueFalse[,names(aggregateResults), drop = FALSE]
  aggregatedTrueNames <- names(aggregatedTrue)
  # add True results to newRecord
  newRecord[1, aggregatedTrueNames] <- aggregateResults[1, aggregatedTrueNames]
  
  # run keepHighestVal() on specific remaining columns
  keptHighest <- oneUniqueFalse %>%
    summarize(across(any_of(keepHighestVars), ~ max(.x, na.rm = TRUE)))
  keptHighestTrueNames <- names(keptHighest)
  newRecord[1, keptHighestTrueNames] <- keptHighest[1,]
  if("lastModifiedDate" %in% keptHighestTrueNames){
    newLastModBy <- compareDF[compareDF$lastModifiedDate == keptHighest$lastModifiedDate, "lastModifiedBy"]
    newRecord[1, "lastModifiedBy"] <- newLastModBy[1]
    keptHighestTrueNames <- append(keptHighestTrueNames, "lastModifiedBy")
  }
  
  # check if any columns remain uncombined
  combinedColNames <- c(identicalTrueNames, oneUniqueTrueNames, aggregatedTrueNames, keptHighestTrueNames)
  uncombinedCols <- compareDF %>%
    select(-all_of(combinedColNames))
  uncombinedNames <- names(uncombinedCols)
  
  # address `numberOfNotes` column
  if("numberOfNotes" %in% uncombinedNames){
    newRecord[1, "numberOfNotes"] <- sum(compareDF$numberOfNotes, na.rm = TRUE)
    uncombinedCols <- uncombinedCols %>%
      select(-numberOfNotes)
    uncombinedNames <- names(uncombinedCols)
  }
  
  # create a few new columns
  newRecord <- newRecord %>%
    rename(allEmailAddresses = primaryEmailAddress, allAcctNums = accountNumber) %>%
    mutate(firstAcctNum = ifelse(str_detect(allAcctNums, "[|]") == TRUE, min(as.numeric(strsplit(allAcctNums, split = "[|]")[[1]])), allAcctNums))
  
  results_list <- list("dupRecordsAllCols" = compareDF, "identicalAcrossRecords" = identicalTrue,
                       "oneUniqueAcrossRecords" = oneUniqueTrue, "aggregatedAcrossRecords" = aggregatedTrue,
                      "combinedColumns" = combinedColNames,
                       "remainingColumns" = uncombinedCols, "deduplicatedRecord" = newRecord)
  
  return(results_list)
}
```

```{r}
collapseRecords(indOnlyRecords, "[Constituent Name]")

adminAssts <- bloomConstsClean %>%
  filter(grepl("house.ga.gov", primaryEmailAddress) == TRUE)

dupEmails <- bloomConstsClean %>%
  group_by(primaryEmailAddress) %>%
  summarize(nOccur = n()) %>%
  filter(nOccur != 1) %>%
  pull(primaryEmailAddress)

dupEmailRecords <- bloomConstsClean %>%
  filter(primaryEmailAddress %in% dupEmails)
  
unique(bloomConstsClean$primaryEmailAddress)

oneEmailManyNames <- bloomConstsClean %>%
  group_by(primaryEmailAddress) %>%
  summarize(numNames = n_distinct(fullName)) %>%
  arrange(desc(numNames))

oneNameManyEmails <- bloomConstsClean %>%
  group_by(fullName) %>%
  summarize(numEmails = n_distinct(primaryEmailAddress)) %>%
  arrange(desc(numEmails)) %>%
  filter(numEmails != 1) %>%
  pull(fullName)

bloomConstsClean %>%
  filter(fullName %in% oneNameManyEmails) %>%
  arrange(fullName) %>%
  relocate(fullName, primaryEmailAddress, type, .before = 1) %>%
  filter(type == "individual")
```

## **Segment 1: Individual Account Types Only**

```{r}
newRecordsDF <- data.frame()
partiallyCollapsed <- vector()
uniqueNames <- unique(indOnlyRecords$fullName)
for(name in uniqueNames){
  recordsList <- collapseRecords(indOnlyRecords, name)
  newRecordsDF <- rbind(newRecordsDF, recordsList[["deduplicatedRecord"]])
  if(length(recordsList[["remainingColumns"]]) != 0){
    partiallyCollapsed <- append(partiallyCollapsed, name)
  }
}

uncollapsedRecords <- indOnlyRecords %>%
  filter(fullName %in% partiallyCollapsed)

uncollapsed <- data.frame()
remainingCols <- vector()
for(name in partiallyCollapsed){
  uncollapsedRecords <- collapseRecords(indOnlyRecords, name)
  remainingCols <- append(remainingCols, names(uncollapsedRecords[["remainingColumns"]]))
  uncollapsed <- rbind(uncollapsed, c(name, str_flatten(names(uncollapsedRecords[["remainingColumns"]]), collapse = ", ")))
}

names(uncollapsed) <- c("Name", "RemainingVars")
uncollapsedSummary <- uncollapsed %>%
  arrange(RemainingVars)

table(remainingCols)
```

# Transaction Data:

```{r}
collapseTransactions <- function(full_name){
transCompareDF <- indOnlyTransactions %>%
  filter(fullName == full_name)

transactionDF2 <- transCompareDF
firstTrans <- transactionDF2[,c("firstTransactionAmount", "firstTransactionDate")]
largestTrans <- transactionDF2[,c("largestTransactionAmount", "largestTransactionDate")]
latestTrans <- transactionDF2[,c("latestTransactionAmount", "latestTransactionDate")]
secondTrans <- transactionDF2[,c("secondTransactionAmount", "secondTransactionDate")]
transDFnames <- c("amount", "date")
names(firstTrans) <- transDFnames
names(largestTrans) <- transDFnames
names(latestTrans) <- transDFnames
names(secondTrans) <- transDFnames

allUnqTrans <- rbind(firstTrans, largestTrans, latestTrans, secondTrans) %>%
  arrange(date) %>%
  mutate(amount = round(as.numeric(amount), digits = 2)) %>%
  unique() %>%
  filter(is.na(amount) == FALSE & is.na(date) == FALSE) %>%
  mutate(transNum = seq.int(nrow(.)))

### Fill in final record fields:
finalRecord$firstTransactionDate <- allUnqTrans$date[allUnqTrans$id == 1]
finalRecord$firstTransactionAmount <- pull(filter(allUnqTrans, date == finalRecord$firstTransactionDate), amount)
finalRecord$secondTransactionDate <- allUnqTrans$date[allUnqTrans$id == 2]
finalRecord$secondTransactionAmount <- pull(filter(allUnqTrans, date == finalRecord$secondTransactionDate), amount)
finalRecord$latestTransactionDate <- max(allUnqTrans$date, na.rm = TRUE)
finalRecord$latestTransactionAmount <- pull(filter(allUnqTrans, date == finalRecord$latestTransactionDate), amount)
finalRecord$largestTransactionAmount <- max(allUnqTrans$amount, na.rm = TRUE)
finalRecord$largestTransactionDate <- pull(filter(allUnqTrans, amount == finalRecord$largestTransactionAmount), date)
finalRecord$numberOfTransactions <- nrow(allUnqTrans)

### Fill in transaction records based on previous transaction vars:
finalRecord$lifetimeRaised <- sum(allUnqTrans$amount)
### We're gonna define "last year" as the
fiscalYearStart <- as.Date("07/01/2020", format = c("%m/%d/%Y"))
fiscalYearEnd <- as.Date("06/30/2021", format = c("%m/%d/%Y"))
finalRecord$lastYearRaised <- allUnqTrans %>%
  filter(date >= fiscalYearStart & date <= fiscalYearEnd) %>%
  summarize(lastYearRaised = sum(amount)) %>%
  pull(lastYearRaised)
}
```

```{r}
newRecordsDF <- data.frame()
partiallyCollapsed <- vector()
uniqueNames <- unique(indOnlyRecords$fullName)
for(name in uniqueNames){
  recordsList <- collapseRecords(indOnlyPersonalInfo, name)
  newRecordsDF <- rbind(newRecordsDF, recordsList[["deduplicatedRecord"]])
  if(length(recordsList[["remainingColumns"]]) != 0){
    partiallyCollapsed <- append(partiallyCollapsed, name)
  }
}
```
